\section{DiFX and pulsars} \label{sec:pulsars}


%TODO:
%profile mode
%  - cormac says > 10 min is problematic
%polyco:
%  - obs code had to be numeric prior to 2.5
%  - always expects earth center based polycos
%tempo2 example (Jan)
%  - discourage use of temp2
%profile2binconfig
%  - 2 modes
%  - single input file allowed
%makeemptybinconfig (JW -- to be added)



DiFX supports three pulsar processing modes:

\begin{itemize}

\item \label{psrmode:binarygate} {\bf Binary Gating}  
A simple on-off pulse accumulation window can be specified with an ``on'' phase and an ``off'' phase.
This can be used to boost the signal to noise ratio of pulsar observations by a factor of typically 3 to 6 and can also be used to search for off-pulse emission.

\item \label{psrmode:matchedgate} {\bf Matched-filter Gating}
If the pulse profile at the observation frequency is well understood and the pulse phase is very well predicted by the provided pulse ephemeris, additional signal to noise over binary gating can be attained by appropriately scaling correlation coefficients as a function of pulse phase.
Depending on the pulse shape, addition gains by a factor of up to 1.5 in sensitivity over binary gating are realizable.

\item \label{psrmode:bin} {\bf Pulsar Binning}
Pulsar {\em binning} is supported within DiFX.
This entails generating a separate visibility spectrum for each requested range of pulse phase.
There are no explicit limits to the number of pulse phase bins that are supported, however, data rates can become increasingly large.
Currently AIPS does not support databases with multiple phase bins. 
Until there is proper post-processig support for pulsar binning, a separate FITS file will be produced for each pulsar phase bin.

\end{itemize}

In all cases the observer will be responsible for providing a pulsar
spin ephemeris.  Except for certain applications of mode 
\ref{psrmode:bin} the ephemeris must be capable of pedicting the absolute
rotation phase of the pulsar.
Enabling pulsar modes incurs a minimum correlation-time penalty of
about 50\%.  High output data rates (computed from time resolution,
number of spectral channels, and number of pulsar bins) may require 
greater correlator resourse allocations.
The details of pulsar observing, including practical details of using
the pulsar modes and limitations imposed by operations, 
are documented at \url{http://library.nrao.edu/public/memos/vlba/up/VLBASU_32.pdf}.

\subsection{Pulse ephemeris}

The use of any pulsar mode requires a pulse ephemeris to be provided by the astronomer.  
This is a table of one or more polynomial entries, each of which evaluates the pulsar's rotation phase over an interval of typically a few hours.
The classic pulsar program {\tt Tempo} can be used to produce the polynomials required \cite{tempo}.
The pulse phase must be evaluated at the Earth center which is usually specified in {\tt tempo} by station code 0 (zero).
Many pulsars exhibit a great degree of timing noise and hence the prediction of absolute pulse phase may require updated timing observations.
When submitting the polynomial for use at the correlator, please adhere to the following naming convention: {\em experiment}{\tt -}{\em pulsar}{\tt .polyco} , e.g., {\tt BB118A-B0950+08.polyco} .
Instructions for generating the polynomial file are beyond the scope of this document.

\subsection{Bin configuration file}

All three pulsar modes also require the preparation of a {\tt .binconfig} file by the astronomer.
The contents of this file determine which of the three pulsar modes is being used.
Three pieces of information are contained within this file: the pulsar ephemeris (polyco) files to apply, definitions of the pulsar bins, and a boolean flag that determines whether the bins are weighted and added within the correlator.
The file consists of a set of keywords (including a colon at the end) that must be space padded to fill the first 20 columns of the file and the values to assign
to these keywords that start at column 21.
The file is case sensitive.
The pulsar bins each consist of a ending phase and a weight; each bin is implicitly assumed to start when the previous ends and the first bin starts at the end phase of the last.
The phases are represented by a value between 0 and 1 and each successive bin must have a larger ending phase than the previous.
Examples for each of the three pulsar modes are shown below:

\subsubsection{Binary gating}

\begin{verbatim}
NUM POLYCO FILES:   1
POLYCO FILE 0:      BB118A-B0950+08.polyco
NUM PULSAR BINS:    2
SCRUNCH OUTPUT:     TRUE
BIN PHASE END 0:    0.030000
BIN WEIGHT 0:       1.0
BIN PHASE END 1:    0.990000
BIN WEIGHT 1:       0.0
\end{verbatim}

\subsubsection{Matched-filter gating}

\begin{verbatim}
NUM POLYCO FILES:   1
POLYCO FILE 0:      BB118A-B0950+08.polyco
NUM PULSAR BINS:    6
SCRUNCH OUTPUT:     TRUE
BIN PHASE END 0:    0.010000
BIN WEIGHT 0:       1.0
BIN PHASE END 1:    0.030000
BIN WEIGHT 1:       0.62
BIN PHASE END 2:    0.050000
BIN WEIGHT 2:       0.21
BIN PHASE END 3:    0.950000
BIN WEIGHT 3:       0.0
BIN PHASE END 4:    0.970000
BIN WEIGHT 4:       0.12
BIN PHASE END 5:    0.990000
BIN WEIGHT 5:       0.34
\end{verbatim}

Note here that there is zero weight given to pulse phases ranging between 0.05 and 0.95.

\subsubsection{Pulsar binning}

\begin{verbatim}
NUM POLYCO FILES:   1
POLYCO FILE 0:      BB118A-B0950+08.polyco
NUM PULSAR BINS:    20
SCRUNCH OUTPUT:     FALSE
BIN PHASE END 0:    0.025000
BIN WEIGHT 0:       1.0
BIN PHASE END 1:    0.075000
BIN WEIGHT 1:       1.0
BIN PHASE END 2:    0.125000
BIN WEIGHT 2:       1.0
BIN PHASE END 3:    0.175000
BIN WEIGHT 3:       1.0
.
.
.
BIN PHASE END 18:   0.925000
BIN WEIGHT 18:      1.0
BIN PHASE END 19:   0.975000
BIN WEIGHT 19:      1.0
\end{verbatim}

The primary difference is {\tt SCRUNCH OUTPUT:     FALSE} which causes each pulsar bin to be written to disk.

\subsection{Preparing correlator jobs}

When using {\tt vex2difx} to prepare correlator jobs, one must associate the pulsar with a setup of its own that includes reference to the {\tt .binconfig} file.
An excerpt from a {\tt .v2d} file is below:

\begin{verbatim}
SETUP gateB0950+08
{
        tInt = 2.000
        nChan = 32
        doPolar = True
        binConfig = BB118A-B0950+08.binconfig
}

RULE B0950+08
{
        source = B0950+08
        setup = gateB0950+08
}
\end{verbatim}

The {\tt .binconfig} file should be in the same path as the {\tt .v2d} file when running {\tt vex2difx}.

\subsection{Making FITS files}

For the two gating modes, preparing FITS files with {\tt difx2fits} is no different than for any other DiFX output.
FITS-IDI does not support multiple phase bins so the pulsar binning case is different and the situation is non-optimal.
Each pulsar bin must be made into its own {\tt FITS} file with a separate execution of {\tt difx2fits}.  
The {\tt -B} (or {\tt --bin}) command line option takes the bin number (starting at zero as above) and writes a FITS file containing data only associated with that bin number.
Be sure to systematically name output files such that the bin number is understood.
